DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-07-20 2015CompleteDerivation/completeDerivation.tex
Final cotangent bundle proof and other improvements

\subsection{Homogeneous decomposable systems}
The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its parts. For example, the state of a ball is equal to the state of its top and bottom parts.

@@

\subsection{Classical homogeneous decomposable systems}
We also want to be able to express the state of the composite system in terms of the state of each part. For example, given the state of a fluid we'll want to know how the material is distributed in space. This in general will depend on how much the state of the composite system "knows" about its parts. For example, the position and orientation of an ideal rigid body is enough to define where all its constituents are, while the volume/pressure/temperature of an ideal gas is not enough to determine the position and momentum of all its constituents. Therefore we need to characterize the system further.

@@

\subsection{Invariant densities}

@@

On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$. For states identified by a single continuous state variable we would expect $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$. By a simple change of state variable, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$. This would make $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using. This is inconsistent with our previous definition. Moreover if the transformation is not differential, the density is not even well-defined. This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.

The first order of business is to guarantee that the density remains defined under an arbitrary change of state variables. This means the Jacobian of the transformation must exist. For the Jacobian to exist the transformation between state variables has to be differentiable. Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold. $\mathcal{S}$ has a differentiable structure in the sense that the distributions and the volume element $d\mu$ are only defined on a set of state variables that are linked by differential transformations. In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differentiable changes of state variable can be detected because they are incompatible with state-variable-independent densities.

@@

\begin{prop}\label{differentiable_manifold}
	The state space $\mathcal{S}$ for the particles of a homogeneous classical material identified by a finite number of continuous independent state variables is a differentiable manifold. The state space $\mathcal{C}$ for a homogeneous classical material is isomorphic to the space of Lebesgue integrable differentiable functions. That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is differentiable$\}$.
\end{prop}
\begin{justification}
	We claim $\mathcal{S}$ is a differential manifold. Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$. $\rho_\mathcal{c}$ is defined on any chart of $\mathcal{S}$. Given two charts, for any $\rho_\mathcal{c}(q)$ there is a corresponding $\rho_\mathcal{c}(q')=\rho_\mathcal{c}(q^i) | d q' / d q |$. The Jacobian $| d q' / d q |$ exists and is non-zero. The map between any two charts is differentiable. $\mathcal{S}$ is a differentiable manifold.

@@

Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the pseudo-norm $\int_{\mathcal{S}} \rho d\mu$ is perhaps more physically meaningful. For proper states, the two are the same and they represent the total amount of material. For state changes they differ. The second represents the amount of material added (or taken away if negative) by the state change. For example, if $\delta \rho$ is the change due to deterministic and reversible evolution, the amount of material does not change and therefore $\int_{\mathcal{S}} \delta \rho \, d\mu = 0$. The first norm would represent the total amount of material that is changing (i.e. being added and being removed), therefore $\int_{\mathcal{S}} | \delta \rho | \, d\mu = 0$ means no material is moving, no change is occurring.

Whatever norm one may choose, $\mathcal{C}$ is a normed vector space but not a complete one. It does not include the limit of all Cauchy sequences as discontinuous functions are excluded. Therefore $\mathcal{C}$ cannot be a Banach or a Hilbert space. As we mentioned before, we can extend the space for computation purposes or to study a limit case as long as we are mindful of the extension and carefully check that the mathematical results do not depend on such extension and are physically meaningful.

@@

If $\rho_\mathcal{c}$ is to remain invariant under state variable changes, not only the Jacobian has to exist but must be unitary. This means we cannot simply change one state variable as we please. If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary. This means that we cannot change physical units of all the variables as we like: they are part of a unit system. As we know, changing units of position will also change units of momentum. Let $\mathcal{Q}$ be the manifold charted by the minimum set of state variables that define such system of units. An element of $\mathcal{S}$ will be a geometrical object in such space. For example, it will be a point, a distance, two points, the ratio between two distances, the inverse of a distance: something that is expressible in the unit defined. As particle states represent configurations of infinitesimal parts, they will be local objects, defined around a point in $\mathcal{Q}$. As we saw before, they cannot be points themselves as the density $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$ is not invariant. They cannot be a vectors $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant. They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$. Physically, the component $k_i$ of our covector, which are the classical analogue of the wave number, can be expressed with a unit that is the inverse of the corresponding $q^i$. If $q^i$ is in meters, $k_i$ is in inverse meters. The product $\Delta q^i \Delta k_i$ is therefore a dimensionless quantity, invariant under unit transformation. A covector is the simplest local geometrical object that allows coordinate independent densities. The cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors, represents the simplest mathematical structure $\mathcal{S}$ can have to be able to support coordinate invariant densities.

If $\mathcal{Q}$ is one dimensional, then the state is fully identified by a pair $(q, k)$. These variables are independent, so the total number of possibilities in a small region of $\mathcal{S}$ is proportional to $dq \wedge dk$. That is if we double the range of $dq$ or $dk$ then the number of possibilities doubles. We set $d\mu = \hbar dq \wedge dk$, where $\hbar$ is the proportionality constant that will depend on the unit chosen to count the possibilities of the pair $(q, k)$. That is, a unitary range of possibilities for $q$ and $k$ will give us $\hbar$ possibilities.\footnote{The actual value and physical dimensions of $\hbar$ are determined by the system of units, and should not be taken to describe some intrinsic physical property. Only dimensionless relationships to other physical constants, such as the fine structure constant $\alpha = e^2/\hbar c$, can possess that trait. That is why one can choose "natural units" for which $\hbar=1$. In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The cardinality for states in a finite (i.e. compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$.

If $\mathcal{Q}$ is $n$-dimensional, $d\mu$ is proportional to $dq^n \wedge dk_n$. But it's not just the volume that is preserved as we change state variables. Each pair $(q^i, k_i)$ represents an independent set of state variables, with the cardinality of possibilities given by $\hbar dq^i \wedge dk_i$. We call such a pair a degree of freedom. As they are independent the total number of states is the product of the possibilities of each d.o.f.: $d\mu = \hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$. This is equivalent to saying that independent d.o.f. are orthogonal surfaces in $\mathcal{S}$.

@@

We can capture the above discussion by stating that the state space $\mathcal{S}$ is the symplectic manifold $(T^*\mathcal{Q}, \omega)$. This is the simplest manifold that allows us to define state-variable-invariant densities. The metric allows us to measure the cardinality of possibilities within and across independent d.o.f.

@@

It should be clear by now that the case of discrete topology is qualitatively different from the standard topology for $\mathbbm{R}^n$. The notion that the continuous case is a limit of the discrete case, that it's ``like the discrete but with more points", leads in this case to erroneous intuition. The question is: can we identify a single state? Can we have an outcome associated with only one element? The answer is yes in the discrete topology and is no in the continuous case. The consequence is that when we define the measure $\mu$ for the cardinality of continuous states, we assign a finite value to a finite region, and we assign zero to a single state. That is what gave us integration, densities and, ultimately, conjugate variables. If we were to use the discrete topology on $\mathbbm{R}^n$, if we were to assume we can identify single states, then we would assign measure one to each state and infinity to a finite region. We would not have integration, just a simple sum. Our distribution would not be a density and there would be no justification for conjugate variables. A finite distribution could only distribute finite amount of material in a finite number of states.

In other words: this is not a case were the difference at the small scale has a small difference in the large scale. The two cases are radically different. We can tell them apart. We should take the use of densities and conjugate quantities as empirical evidence that the quantities like space and time are not discrete. Or more precisely: that the processes we use to distinguish those quantities cannot identify single instances, as that is all we can say experimentally. The quantum case does not change this, as the use of densities (in the form of the wave function) and conjugate quantities are even more prominent.

Finally, we should note that we in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states. We have both a topological space and a metric that allows us to count states and possibilities. We recovered the metric starting from the idea of an infinitesimally reducible system. That led to a state space $\mathcal{S}$ for the infinitesimal parts, on which we must be able to define state-variable-independent distributions, which in turn gave us degrees of freedom made by pairs of conjugate variables and the symplectic metric typical of classical phase space. In short: being able to measure the amount of material is what allows us to measure the cardinality of states.

@@

Now that we have fully characterized what we mean by a classical system, we can stipulate the following:

@@

	In other words: the definition of a state requires some kind of asymmetry between system under study and environment. If we try to assume that both the system and environment are really made of infinitesimal homogeneous parts, then the claim to that asymmetry is lost. Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale but large compared to another. For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules.

	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we do not want the system isolated so that we can define physical processes on it. For example, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts. Each particle would essentially reside in its own separate physical universe.

	If we just assume the whole system is deterministic and reversible, then each part can evolve depending on the states of the other parts. But this does not solve the problem entirely. Let's ignore the question of how we were able to define the states of the parts. Let's also ignore that the type of forces one would typically use (i.e. ones that conserve energy or ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other. The problem is that we can always locally separate the evolution into independent degrees of freedom. For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently. And since the pieces are infinitesimal, a local decomposition is all that is needed. While such description may be cumbersome to achieve in practice, conceptually it is still possible. So, even if the particles are in principle interacting, the system is described by degrees of freedom that evolve independently, oblivious to each other.

@@

@@

	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to apply to all of physics. While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and is also useful as a default assumption of sorts. Understanding its shortcomings will help us later to see how the quantum assumption solves, at least partially, some of these issues.

>>>>

\subsection{Homogeneous decomposable systems and vector spaces}
The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its parts. For example, the state of a ball is equal to the state of its top and bottom parts.

@@

\subsection{Classical material and distributions}
We also want to be able to express the state of the composite system in terms of the state of each part. For example, given the state of a fluid we'll want to know how the material is distributed in space. This in general will depend on how much the state of the composite system "knows" about its parts. For example, the position and orientation of an ideal rigid body is enough to define where all its constituents are, while the volume/pressure/temperature of an ideal gas is not enough to determine the position and momentum of all its constituents. Therefore we need to characterize the system further.

@@

\subsection{Integration and measure}

@@

\subsection{Invariant densities and differentiability}

On a state space with discrete topology, using a single state variable, we have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$. We would expect the expression to become $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$ for states identified by a continuous state variable. By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$. This makes $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using. This is inconsistent with our previous definition. Moreover if the transformation is not differential, the density is not even well-defined. This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.

The first order of business is to guarantee that the density remains defined under an arbitrary change of state variables. This means the Jacobian of the transformation must exist. For the Jacobian to exist the transformation between state variables has to be differentiable. Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold. $\mathcal{S}$ has a differentiable structure in the sense that the distributions and the volume element $d\mu$ are only defined on a set of state variables that are linked by differential transformations. In the same way that discontinuous changes of state variable can be detected because they tamper with physical distinguishability (i.e. the topology), non differentiable changes of state variable can be detected because they tamper with our ability to define state-variable-independent densities.

@@

\begin{prop}\label{differentiable_manifold}
	The state space $\mathcal{S}$ for the particles of a  classical material identified by a finite number of continuous independent state variables is a differentiable manifold. The state space $\mathcal{C}$ for a homogeneous classical material is isomorphic to the space of Lebesgue integrable differentiable functions. That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is differentiable$\}$.
\end{prop}
\begin{justification}
	We claim $\mathcal{S}$ is a differential manifold. Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$. $\rho_\mathcal{c}$ is defined on any chart of $\mathcal{S}$. Given two charts, for any $\rho_\mathcal{c}(q)$ there is a corresponding $\rho_\mathcal{c}(q')=\rho_\mathcal{c}(q^i) | d q' / d q |$. The Jacobian $| d q' / d q |$ exists and is non-zero. The map between any two charts is differentiable. $\mathcal{S}$ is a differentiable manifold.

@@

Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the expression $\int_{\mathcal{S}} \rho d\mu$ is perhaps more physically meaningful. For proper states, the two are the same and they represent the total amount of material. For state changes they differ. The second represents the amount of material added (or taken away if negative) by the state change. For example, if $\delta \rho$ is the change due to deterministic and reversible evolution, the amount of material does not change and therefore $\int_{\mathcal{S}} \delta \rho \, d\mu = 0$. The first norm would represent the total amount of material that is changing (i.e. being added and being removed), therefore $\int_{\mathcal{S}} | \delta \rho | \, d\mu = 0$ means no material is moving, no change is occurring.

Note that neither of those expressions match the norm given by the vector product defined as $\langle f, g \rangle = \int_{\mathcal{S}} fg \, d\mu$. Moreover, since $\mathcal{C}$ excludes discontinuous functions, it can never be a complete normed space: it will not include the limit for all Cauchy sequences. Such constructions, though, are still useful. Consider the following expressions:
\begin{align*}
\Lambda_{\mathcal{S}} (\mathcal{c}) &= \int_\mathcal{S} \rho_{\mathcal{c}} d \mu = \int_\mathcal{S} 1\cdot  \rho_{\mathcal{c}} d \mu \\
\Lambda_U (\mathcal{c}) &= \int_U \rho_{\mathcal{c}} d \mu = \int_\mathcal{S} 1_U \cdot \rho_{\mathcal{c}} d \mu \\
p_{total}(\mathcal{c}) &= \int_\mathcal{S} p \cdot \rho_{\mathcal{c}} d \mu \\
q_{avg}(\mathcal{c}) &= 1/|\rho_{\mathcal{c}}| \int_\mathcal{S} q \cdot \rho_{\mathcal{c}} d \mu
\end{align*}
For each state, they represent respectively the total amount of material, the material within $U$, the total momentum and the average of a state variable. As they are linear functionals of $\mathcal{C}$,  they can be expressed using their dual vector over $\mathcal{S}$. But these vectors ($1$, $1_U$, $p$, $q/|\rho_{\mathcal{c}}|$) are \emph{not} elements of $\mathcal{C}$. They do not represent states, they are not continuous and integrable, and there is no single general physical meaning for them. Therefore, as we mentioned before, while we can extend the function space for convenience, computation purposes or to study a limit case, we need to be mindful of the extension and carefully check that the mathematical results are physically meaningful case by case.

\subsection{Invariant densities and cotangent bundle}

@@

If $\rho_\mathcal{c}$ is to remain invariant under state variable changes, not only the Jacobian has to exist but must be unitary. This means we cannot simply change one state variable as we please. If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary. This  means that we cannot change physical units of all the variables as we like: they are part of a unit system.

The simplest case is when a single variable is enough to define our units, and therefore tell how all the others must change. Suppose that our state space $\mathcal{S}$ is identified by $n$ independent state variables $\{q, k^i\}$ and a change of units for the first, that is $\hat{q} = \hat{q}(q)$, determines the change for all others. The Jacobian matrix is the block matrix:
\begin{align*}
J =  \left[
\begin{array}{cc}
d_q\hat{q} & 0 \\
\partial_q \hat{k}^j & \partial_i \hat{k}^j \\
\end{array}
\right] 
\end{align*}
The Jacobian has to be unitary, therefore all elements $\partial_i \hat{k}^j$ must be well defined and $|\partial_i \hat{k}^j| = 1/d\hat{q}$. This only happens if $\partial_i \hat{k}^j$ is a $1\times1$ matrix: there is only one $k$. This gives means $\partial_{k} \hat{k} = d_{\hat{q}} q$ and $\hat{k} = d_{\hat{q}} q \, k + f(q)$. As this is a change of units, we expect the zero value of $k$ to remain constant: $\hat{k}(q, 0) = 0 = f(q)$. Therefore $\hat{k} = d_{\hat{q}} q \, k$ transforms as the component of a covariant vector. $\mathcal{S}$ is isomorphic to $T^*\mathcal{Q}$ where $\mathcal{Q}$ is the manifold that defines our unit.

Physically, this means that the state variable $k$, which is the classical analogue of the wave number, is expressed with the inverse of the unit used for $q$. If $q$ is in meters, $k$ is in inverse meters. Consider now the area $dq \wedge dk$ of an infinitesimal rectangular region. This quantity is dimensionless and therefore invariant. The total number of states will be proportional to it, as  doubling the range of $dq$ or $dk$ will give us double the number of possible states. We have $d\mu = \hbar dq \wedge dk$, where $\hbar$ is the proportionality constant that will depend on the unit chosen to count the possibilities of the pair $(q, k)$. That is, a unitary range of possibilities for $q$ and $k$ will give us $\hbar$ possibilities.\footnote{The actual value and physical dimensions of $\hbar$ are determined by the system of units, and should not be taken to describe some intrinsic physical property. Only dimensionless relationships to other physical constants, such as the fine structure constant $\alpha = e^2/\hbar c$, can possess that trait. That is why one can choose "natural units" for which $\hbar=1$. In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The cardinality for states in a finite (i.e. compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$. We are finally able to write the volume of integration in terms for state variables.

We can generalize to the case where $\mathcal{Q}$ has more than one dimension. As we must be able to change one unit at a time, independently from the other, to each $q^i$ will correspond a $k_i$ that uses the inverse of the corresponding units. We have $\mathcal{S}\cong T^*\mathcal{Q}$ and $d\mu = \hbar^n dq^n \wedge dk_n$. But it's not just the volume that is preserved as we change state variables. Each pair $(q^i, k_i)$ is expressed by an independent unit with the cardinality of possibilities given by $\hbar dq^i \wedge dk_i$. We call such a pair a degree of freedom. As they are independent the total number of states is the product of the possibilities of each d.o.f.: $d\mu = \hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$. In other words: independent d.o.f. are orthogonal surfaces in $\mathcal{S}$. The possibilities of an arbitrary degree of freedom, then, will be the sum of the projections over the $n$ orthogonal and independent d.o.f. defined by $(q^i, k_i)$.

@@

We can capture the above discussion by stating that the state space $\mathcal{S}$ is the symplectic manifold $(T^*\mathcal{Q}, \omega)$, where $\mathcal{Q}$ is the manifold that defines the unit system. This is the only manifold that allows us to define state-variable-invariant densities. The metric allows us to measure the cardinality of possibilities within and across independent d.o.f.

@@

It should be clear by now that the case of discrete topology is qualitatively different from the standard topology for $\mathbbm{R}^n$. The notion that the continuous case is a limit of the discrete case, that it's ``like the discrete but with more points", leads in this case to erroneous intuition. The question is: can we identify a single state? Can we have an outcome associated with only one element? The answer is yes in the discrete topology and is no in the continuous case. The consequence is that when we define the measure $\mu$ for the cardinality of continuous states, we assign a finite value to a finite region, and we assign zero measure to a single state. That is what gave us integration, densities and, ultimately, conjugate variables. If we were to use the discrete topology on $\mathbbm{R}^n$, if we were to assume we can identify single states, then we would assign measure one to each state and infinity to a finite region. We would not have integration, just a simple sum. Our distribution would not be a density and there would be no justification for conjugate variables. A finite distribution could only distribute finite amount of material in a finite number of states.

In other words: this is not a case were a difference at  small scale has a small difference at large scale. The two cases are radically different. We can tell them apart. We should take the use of densities and conjugate quantities as empirical evidence that the quantities like space and time are not discrete. Or more precisely: that the processes we use to distinguish those quantities cannot identify single instances, as that is all we can say experimentally. The quantum case does not change this, as the use of densities (in the form of the wave function) and conjugate quantities are even more prominent.

Finally, we should note that we in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states. We have both a topological space and a measure expressible in terms of state variables that allows us to count states and possibilities. We recovered the metric starting from the idea of an infinitesimally reducible system. That led to a state space $\mathcal{S}$ for the infinitesimal parts, on which we must be able to define state-variable-independent distributions, which in turn gave us degrees of freedom made by pairs of conjugate variables and the symplectic metric typical of classical phase space. In short: being able to measure the amount of material is what allows us to measure the cardinality of states.

@@

Now that we have fully characterized what we mean by a classical material, we can stipulate the following:

@@

	In other words: the definition of a state requires some kind of asymmetry between system under study and environment. The environment should not affect the system, so that we can define an independent state, yet the system must affect the environment, so that we have external process that allows to define physical distinguishability. If we try to assume that both the system and environment are really made of infinitesimal homogeneous parts, then the claim to that asymmetry is lost. Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale (they can be regarded as infinitesimal) but large compared to another (they are little affected by the environment). For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules such that the average system is sufficiently stable.

	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we want the environment to not be isolated so that we can define physical processes that interact with the system. The problem is that the same physical object could be either under study or part of the environment depending on the situation. In other words: we want systems to be interacting so that they can be physically aware of each other, yet we want them not interacting so that we can properly define states on them. We clearly can't have it both ways. Under our classical assumption, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts. Each particle would essentially reside in its own separate physical universe, as it cannot physically distinguish anything else. This cannot work.

	If we assumed the system is deterministic and reversible only as a whole, then each part could evolve depending on the states of the other parts. This would seem much better, as we have pieces with no unstated part that are physically aware of each others. But this does not actually solve the problem of independent oblivious components. Let's ignore the question of how we were able to define the states of the parts. Let's also ignore that the type of forces one would typically use (i.e. ones that conserve energy or ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other. The actual problem is that we can always locally separate the evolution into independent degrees of freedom. For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently. And since the pieces are infinitesimal, a local decomposition is all that is needed. While such description may be cumbersome to achieve in practice, conceptually it is still possible. So, even if the particles are in principle interacting, the system is described by degrees of freedom that evolve independently, oblivious to each other.

@@

	The message is that we should \emph{always} beware of thinking that the state of a system is its full description: there is always some unstated part that allows continuous interaction between the system and the environment. That is the part that cannot be isolated (i.e. we can't take the system out of the physical universe) and therefore can't be assigned a state.

@@

	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to strictly apply to all of universe. While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and is also useful as a default assumption of sorts. Understanding its shortcomings will help us later to see how the quantum assumption solves, at least partially, some of these issues.
